Real-Time Location Recommendation for Field Data Collection

### 2017-02-26 Dynamic Enumerator Allocation/Paper/v1/Paper.tex
Add paper v1

% *** MATH PACKAGES ***
%
\usepackage{amsmath}
% A popular package from the American Mathematical Society that provides
% many useful and powerful commands for dealing with mathematics.
%
% Note that the amsmath package sets \interdisplaylinepenalty to 10000
% thus preventing page breaks from occurring within multiline equations. Use:
%\interdisplaylinepenalty=2500
% after loading amsmath to restore such page breaks as IEEEtran.cls normally
% does. amsmath.sty is already installed on most LaTeX systems. The latest
% version and documentation can be obtained at:
% http://www.ctan.org/pkg/amsmath

@@

@@

\begin{document}
%
% paper title
% Titles are generally capitalized except for words such as a, an, and, as,
% at, but, by, for, in, nor, of, on, or, the, to and up, which are usually
% not capitalized unless they are the first or last word of the title.
% Linebreaks \\ can be used within to get better formatting as desired.
% Do not put math or special symbols in the title.
\title{Design and Implementation Mobile Data Collection\\Using Mobile Proxy}

@@

% As a general rule, do not put math, special symbols or citations
% in the abstract
\begin{abstract}
Data collection using mobile device is something that is quite common recently. However, most of those applications are using static validation rules. The use of mobile proxy to duplicate rule and data is believed could improve the flexibility and implementation of the validation rule. Rule could be easily designed in modular basis, installed, updated, as well as deleted from application using OSGi framework. This paper is aimed to research about the design and implementation of mobile proxy that could be used to implement dynamic validation rule on SOA-based mobile device.
\end{abstract}

@@

\section{Introduction}
Data collection is one of the main tasks of statistical institution in the world, including BPS-Statistics Indonesia. Data collection process is conducted using various media, ranging from conventional method, such as paper questionnaire to the modern method using mobile device. Until now, most of data collection conducted by BPS-Statistics Indonesia are still using paper questionnaire. However, BPS is now starting using mobile device for collecting data.

In data collection, the most important thing to be considered is the data quality. Data quality is measured using consistency and validation parameter. Data collection using paper questionnaire still rely on human’s carefulness and thoroughness in implementing the rule. Using mobile device has a benefit related to consistency and validation. Thus, generally, using mobile device could increase data quality.

The scenario that is commonly used in the implementation of mobile device application is to wrap it in a package including GUI and workflow. This scenario has a weakness, that workflows are used are static and cannot be added, removed, or updated, except by updating the whole application.

Another scenario which is also commonly used is by adopting a Service-Oriented approach. Service oriented approach basically separated into two components: the service provider (server) and a service consumer (client). In service-based approach, the workflow will be executed on the server, while the client access the workflow through a network. Service-based approach has been widely implemented, including implementation on mobile based applications. However, the implementation of service-based approach to the mobile device has several potential obstacles, mainly due to the availability of the Internet network, so mobile device that acts as a client cannot always connect to the server. No connection of mobile devices on the server can be categorized into three types \cite{gutwin_gone_2010}:

\begin{itemize}
  \item \textbf{Delay-based interruption:} delays in message delivery because of problems that occur between the sender and receiver, 
  \item \textbf{Network outage:} a condition where a node is disconnected from the other nodes. Cause outage network include: network failure, mobile devices moving between zones, laptops going to sleep, unplugged cables, or inadvertent closing of the application,
  \item \textbf{Explicit departure:} a condition in which the user explicitly logs out of the session or quits the application.
\end{itemize}

The proposed alternative scenario is to adopt a proxy approach. Proxy approach basically is to put a server between the service provider and the service consumer that will intercept the request from the client to the server. Proxy can act as a forward proxy in a condition connected to the network, or cache server in conditions not connected with the network. Illustration of proxy will be designed can be seen in \figurename{\ref{fig:proxy-overview}}.

\begin{figure}
    \centering
    \includegraphics[width=8.5cm]{../../Resources/Images/proxy-overview}
    \caption{Proxy Overview}
    \label{fig:proxy-overview}
\end{figure}

This paper will examine how to design a proxy that can be used in a condition connected or disconnected, which is divided into five sections. Section 1 is the background of this research. While the second section will discuss the literature review and research related to the design of the proxy. Section 3 will discuss the proposed design, followed by a discussion in section 4. Finally section 5 concludes the paper with a summary and a discussion of future work.

\section{Theoretical Background}
In general, the proxy can be said as a placeholder for other entities \cite{gamma_design_1995}. Due to its capability to intercept communication source and target, it allows deferring the initialization of the target, redirecting the communication target to one or more entities, performing additional capabilities before/after communicating with the target, etc. \cite{gani_improving_2009}.

Implementation of the proxy have been conducted with various purposes, e.g., for memory efficiency \cite{gamma_design_1995}, remote communication \cite{wilson_get_2000}, and concurrent data evaluation/processing (e.g. for future object) \cite{pratikakis_transparent_2004}. Proxies have also been used in component-based middleware such as EJB \cite{_fusion_????}, to provide various enterprise-level capabilities. Furthermore, proxies have also been widely adopted in contemporary work on application adaptation \cite{holder_system_1999} \cite{philippsen_javaparty_1997} \cite{rubinsztejn_framework_2005} \cite{ryan_application_2004} \cite{tatsubori_bytecode_2001} \cite{tilevich_j-orchestra:_2002}, since they allow certain components of an application to dynamically reconfigure themselves (e.g. changing its behavior) transparently, i.e. without the awareness of the rest of the components in the application. While on the mobile device, the proxy is used to bridge the wired-wireless gap, and make all mobility, connectivity and context-dependent issues transparent to the application developer \cite{rubinsztejn_framework_2005}.

To support the mobility of a mobile device, the proxy must also be portable and dynamic. Rubinsztejn et al make the architectural design of the proxy in the form of middleware, which consists of both server and client API \cite{rubinsztejn_framework_2005} called MoCA. MoCA architecture consists of a monitor and several services. Monitor is a daemon or service that runs on each mobile client to collect and submit data to the services that run on the server side.

Another approach in the design of proxy proposed by Cobarzan et al. To support the dynamism of proxy, they composed proxy that consists of two components, the daemon and the dispatchers \cite{cobarzan_dynamic_2005}. Dispatcher is a component that acts as proxy. While Dispatcher is used to handle incoming client request, then forward them to the server or to other dispatchers, process it itself, or reject it. Meanwhile, the daemon is a component that runs automatically in the background and is used to monitor the code that contains the command to enable dispatchers. So that the proxy is not automatically activated by the daemon is active before, contrasts with the MoCA which automatically activates the proxy.

Proxy has several functions, such as content adaptation, protocol translation, user authentication, handover management, and cache management \cite{rubinsztejn_framework_2005}. Cache has an important role in reducing latency and network traffic between server and client \cite{rizzo_replacement_2000}. Caching can be implemented in several locations: at the client \cite{bestavros_application-level_1995} \cite{cao_cost-aware_1997}, server \cite{arlitt_trace-driven_1997} \cite{bestavros_application-level_1995}, and within the network \cite{abrams_caching_1995} \cite{cao_cost-aware_1997} \cite{busari_sensitivity_2001} through a proxy server.

Caching is generally implemented on a web the document to store frequently accessed parts of the document \cite{mahanti_temporal_2000} \cite{shim_proxy_1999-1} \cite{busari_sensitivity_2001}, either dynamic or static web contents, although not uncommon also implemented at the application level \cite{holder_system_1999} \cite{philippsen_javaparty_1997} \cite{rubinsztejn_framework_2005}. Since the cache is a temporary object \cite{davison_survey_1999} that represents the data on the server, it needs a method to ensure the consistency of data stored in the cache and which are contained in the server.

Caching method in a distributed system, including on a mobile devices, have a different design with a centralized system. Takdir et al proposed a method of proxy, by caching both data and workflow that will be used by the composite application on the local storage \cite{takdir_multi-layer_2014}. This method provides a transparent resources (i.e. the data and web service logics) access using combination of synchronization, replication, and routing mechanism.

Meanwhile, Terry et al using XML web services caching method \cite{terry_caching_2003} that mimics the behavior of Web services to a limited extent. This XML cache is transparent either on the client side and the server component on a Web services. This method is able to handle disconnections for each Web services. However, this method is lack of consistency, so that an operation performed by the local user could change some of the results of the earlier requests that are stored in the cache.

In a workflow-based application, such as the application of data collection, in which workflows are defined, managed, and executed in accordance with the sequence of logic \cite{wieland_towards_2009}, one of the standard that is now being widely used is the Business Process Execution Language (BPEL) \cite{pautasso_restful_2009} \cite{weerawarana_web_2005}, BPEL is built on the Web Service (WS) standards and Provides a recursive aggregation models for Web services \cite{pasley_how_2005} and used as glue between interacting services \cite{weerawarana_web_2005}. BPEL drives communication of web services in order to deliver business services that Contain several interconnected processes. Web services address interoperability issues in service level, whereas BPEL orchestrate web services in process level. BPEL is ideally suited to the service-oriented architecture, a set of guidelines for integrating disparate systems by presenting each system as a service that implements a specific business function [29].

\begin{figure}
    \centering
    \includegraphics[width=8cm]{../../Resources/Images/ws-standards-stack}
    \caption{Web Service Standards Stack}
    \label{fig:ws-standards-stack}
\end{figure}

\section{Mobile Proxy Design}

\subsection{Design Overview}
Proposed proxy mobile design consists of three components, data, services, and applications, which are connected via two buses, the data bus and bus service. Illustration of design can be seen in \figurename{\ref{fig:proxy-overview}}.

Data bus Provides transparent access to the data source for web services. The data bus consists of two layers, ie cluster database layer and proxy layer. The cluster database layer consists of several database nodes that are equipped with a control cluster to ensure data synchronization. Meanwhile, the proxy layer consists of a proxy server that acts as a load balancer for database nodes. Load balancer is equipped with a healthy check to check the availability of each database node.

Similar with the data bus, the service bus handle web service invocation by the application. Web services that are representation of the rules, are packed as a bundle. Then the bundles are replicated, by composite application, from central repository to local repository. The bundles that are available in local repository will be installed and started dynamically by composite application. Moreover, installed bundles can also be stopped, removed, and updated easily.

\begin{figure}
    \centering
    \includegraphics[width=8cm]{../../Resources/Images/proxy-design-data-layers}
    \caption{Data Bus Layers}
    \label{fig:proxy-design-data-layers}
\end{figure}

\subsection{Replication}
Data and web services bundles are replicated from central storage to local storage as cache objects. Firstly, composite application replicate web services bundles from central repository to local repository. Each web services bundle then will replicate data that it is depend on, from central database into local storage. \figurename{\ref{fig:replication-flowchart}} shows web services and data replication strategy.

\begin{figure}
	\centering
	\subfloat[Web Services Replication Strategy]{
		\centering
		\includegraphics[width=3.5cm]{../../Resources/Images/bundle-replication-flowchart}
		\label{fig:bundle-replication-flowchart}
	}
	\hfil\hfil
	\subfloat[Data Replication Strategy]{
		\centering
		\includegraphics[width=3.5cm]{../../Resources/Images/data-replication-flowchart}
		\label{fig:data-replication-flowchart}}
	\caption{Replication Strategies}
	\label{fig:replication-flowchart}
\end{figure}

\subsection{Synchronization}
Synchronization ensure the data and web service bundles are consistent between local and central repository. Web service bundle must implement versioning, so central server and proxy server must install bundles with same version to ensure the same logic is used. Composite application will periodically execute version check of the bundles, and replicate it when new versions are available in central repository.

Data synchronization is very depend on bundle, since each bundle require different data. Data will be synchronized in two way, from central database to local cache storage (e.g. master data) and from local cache storage to central storage (e.g. transaction data). \figurename{\ref{fig:data-sync-flowchart}} shows data syncronization strategy.

\begin{figure}
    \centering
    \includegraphics[width=8cm]{../../Resources/Images/data-sync-flowchart}
    \caption{Data Syncronization Strategy}
    \label{fig:data-sync-flowchart}
\end{figure}

\subsection{Routing}
Routing plays important role in distributed system. It manages the network traffic and point requests to appropriate destination \cite{takdir_multi-layer_2014}. Local web services are act as proxy server that intermediate composite application, that act as client, and central server. Proxy server will decide whether to forward request to central server or reply it on behalf of central server. \figurename{\ref{fig:routing-flowchart}} shows routing strategy.

\begin{figure}
    \centering
    \includegraphics[width=6cm]{../../Resources/Images/routing-flowchart}
    \caption{Routing Strategy}
    \label{fig:routing-flowchart}
\end{figure}

\section{Implementation}

\subsection{OSGi Framework}
There are several implementations of OSGi, or commonly referred to OSGi Framework, ie: Apache Felix, Concierge, Equinox Eclipse, JBoss, Hitachi, Knopflerfish, and ProSyst. This study uses the Knopflerfish OSGi Framework, because Knopflerfish OSGi Framework has implemented the latest OSGi specification, the OSGi R6. In addition, the Knopflerfish OSGi Framework is also proven to be implemented in Android application.

\subsection{Android Application} \label{ssec:android-application}
By default, Knopflerfish provides an implementation of the OSGi Framework that can be run on any operating system that supports Java SE. However, because the Android OS uses a different JVM implementations, the Dalvik VM, so it needs to be made Knopflerfish OSGi Framework implementations that can run on the Android OS.

Android application that is made must support the principle of OSGi, which is modular. The case studies used in this study is a case study for field data collection, so that the user interface is a questionnaire form. To accelerate the creation and simplify the loading of the page, so it is used Cordova. Cordova is a webview-based, so the control used is an HTML form element. \figurename{\ref{fig:android-app}} shows screenshots of Android application.

\begin{figure}
	\centering
	\subfloat[Home Page]{
		\centering
		\includegraphics[width=2.5cm]{../../Resources/Images/android-app-1}
		\label{fig:android-app-1}
	}
	\hfil
	\subfloat[Autostart OSGi Framework]{
		\centering
		\includegraphics[width=2.5cm]{../../Resources/Images/android-app-2}
		\label{fig:android-app-2}}
	\hfil
	\subfloat[Bundle Page]{
		\centering
		\includegraphics[width=2.5cm]{../../Resources/Images/android-app-3}
		\label{fig:android-app-3}}
	\caption{Android Application}
	\label{fig:android-app}
\end{figure}

\subsection{OSGi Bundles}

\subsubsection{Proxy Server}
Proxy server is an OSGi bundle that implements an embedded web server. Web server used in this study is Jetty 9 which is the highest major version available today. In the Proxy Server Bundle registered a service (API) ContextHandlerService, which is used to register the ServletContext initiated on Application Bundle.

\subsubsection{Composite Application}
Composite Application is an OSGi bundle that implements web services and clients in a same bundle. Web services are implemented in the form of Servlets collected in a single context. While the client is implemented in the form of a web-based application, written in HTML, Javascript, and CSS. In addition, the Composite Application also provides an API that can be imported dynamically by another bundle, ie: AccountHandlerService, DaoHandlerService, and PropertyHandlerService.

\subsubsection{Synchronizer}
Synchronizer is an OSGi bundle that is used to synchronize data in two directions: Remote to Local and Local to Remote. Synchronizer bundle requires some service provided by Composite Application, ie: AccountHandlerService to access the session to be synchronized and DaoHandlerService to access local storage.

\figurename{\ref{fig:android-bundle}} shows screenshots of OSGi bundles installed and started in Android environment.

\begin{figure}
	\centering
	\subfloat[Install bundle]{
		\centering
		\includegraphics[width=2.5cm]{../../Resources/Images/android-bundle-1}
		\label{fig:android-bundle-1}}
	\hfil
	\subfloat[Bundle is installing]{
		\centering
		\includegraphics[width=2.5cm]{../../Resources/Images/android-bundle-2}
		\label{fig:android-bundle-2}}
	\hfil
	\subfloat[Bundles are installed]{
		\centering
		\includegraphics[width=2.5cm]{../../Resources/Images/android-bundle-3}
		\label{fig:android-bundle-3}}
	\vfil
	\subfloat[Bundle is starting]{
		\centering
		\includegraphics[width=2.5cm]{../../Resources/Images/android-bundle-4}
		\label{fig:android-bundle-4}}
	\hfil
	\subfloat[Bundles are active]{
		\centering
		\includegraphics[width=2.5cm]{../../Resources/Images/android-bundle-5}
		\label{fig:android-bundle-5}}
	\hfil
	\subfloat[Composite application is loaded]{
		\centering
		\includegraphics[width=2.5cm]{../../Resources/Images/android-bundle-6}
		\label{fig:android-bundle-6}}
	\caption{Android Bundles}
	\label{fig:android-bundle}
\end{figure}

\subsection{Remote Server}
As described in \ref{ssec:android-application}, Knopflerfish by default already provides an implementation of the OSGi Framework that can be run on any operating system that supports Java SE. Remote server running Knopflerfish using minimal bundle, and installed the same module with the composite application. Listing.\ref{lst:start-framework-remote} and Listing.\ref{lst:start-bundle-remote} shows screenshots of OSGi bundles are active in remote server.

\begin{listing}
    \caption{Knopflerfish with minimal bundles}
    \begin{minted}[showspaces=false,breaklines=true]{java}
java -jar framework.jar -xargs minimal.xargs
Knopflerfish OSGi framework launcher, version <unknown>
Copyright 2003-2015 Knopflerfish. All Rights Reserved.
See http://www.knopflerfish.org for more information.

Created Framework: org.knopflerfish.framework, version=7.2.0.
> Framework launched
	\end{minted}
    \label{lst:start-framework-remote}
\end{listing}

>>>>

% *** MATH PACKAGES ***
%
\usepackage[fleqn]{amsmath}
% A popular package from the American Mathematical Society that provides
% many useful and powerful commands for dealing with mathematics.
%
% Note that the amsmath package sets \interdisplaylinepenalty to 10000
% thus preventing page breaks from occurring within multiline equations. Use:
%\interdisplaylinepenalty=2500
% after loading amsmath to restore such page breaks as IEEEtran.cls normally
% does. amsmath.sty is already installed on most LaTeX systems. The latest
% version and documentation can be obtained at:
% http://www.ctan.org/pkg/amsmath

% *** MATH PACKAGES ***
%
\usepackage{amssymb}

@@

%
% Digunakan untuk membuat inline list dan numbering
\usepackage{paralist}

%
% 
\usepackage{algorithm}
\usepackage{algorithmic}

%
% Digunakan untuk membuat tabel yang panjang, lebih dari 1 halaman
%
\usepackage{longtable, supertabular, booktabs}
  \newcommand{\ra}[1]{\renewcommand{\arraystretch}{#1}}

@@

\begin{document}
%
% paper title
% Titles are generally capitalized except for words such as a, an, and, as,
% at, but, by, for, in, nor, of, on, or, the, to and up, which are usually
% not capitalized unless they are the first or last word of the title.
% Linebreaks \\ can be used within to get better formatting as desired.
% Do not put math or special symbols in the title.
\title{Real-Time Location Recommendation for Field Data Collection}

@@

% As a general rule, do not put math, special symbols or citations
% in the abstract
\begin{abstract}
Field data collection is one of the main activities performed by the statistical agencies of a country. Data collection activities have a common workflow with Multi Depot Vehicle Routing Problem (MDVRP). The use MDVRP to generate pre-calculated routes resulted in a total route costs with high standard deviation. Real-time mechanism by utilizing the paradigm of publish / subscribe combined with COEs MDVRP algorithm is proposed to reduce the inequality (large variation) the time of completion. The test results show that the mechanisms pub / sub combined with COEs produce a total route times more prevalent among enumerators compared with pre-calculated COEs algorithm.
\end{abstract}

@@

%-----------------------------------------------------------------------------%
\section{Introduction}
\label{sec:introduction}
%-----------------------------------------------------------------------------%
Pengumpulan data lapangan merupakan salah satu tugas dan wewenang dari lembaga statistik suatu negara. Terdapat dua macam metode pengumpulan data yang digunakan, yaitu: pengumpulan data primer dan pengumpulan data sekunder. Pada pengumpulan data primer, petugas pengumpulan data secara langsung melakukan interview dengan responden, sementara pada pengumpulan data sekunder, instansi melakukan kompilasi data yang telah dilakukan oleh pihak lain. Pada proses pengumpulan data (primer), petugas pengumpulan data dialokasikan pada sejumlah lokasi pencacahan yang telah ditentukan. Mekanisme pengalokasian yang biasanya digunakan adalah alokasi dengan jumlah lokasi yang sama antar petugas.

Alur kerja pada proses pengumpulan adalah, setiap petugas pengumpulan data memulai pengumpulan data dari sebuah titik tertentu, baik itu kantor maupun tempat tinggal masing-masing petugas. Kemudian petugas melakukan pengumpulan data pada sebuah lokasi sampai selesai, kemudian dilanjutkan dengan lokasi berikutnya, dan seterusnya sampai seluruh lokasi pencacahan dikunjungi. Alur kerja pengumpulan data ini sangat mirip dengan \textit{Vehicle Routing Problem} (VRP), atau lebih spesifik lagi \textit{Multi Depot} VRP.

Akan tetapi, meskipun masalah pengumpulan data memiliki kesamaan alur kerja dengan MDVRP, akan tetapi implementasi algoritma MDVRP secara mentah tidak dapat menjadi solusi atas permasalahan alokasi petugas pengumpulan data, karena salah satu komponen yang penting, yaitu lamanya waktu pelayanan, tidak tersedia sampai dengan lokasi dikunjungi. This research aims to propose an integration antara MDVRP dengan mekanisme pub/sub untuk menciptakan sistem rekomendasi lokasi pencacahan secara real-time.

%-----------------------------------------------------------------------------%
\section{Literature Review}
\label{sec:literature-review}
%-----------------------------------------------------------------------------%
%-----------------------------------------------------------------------------%
\subsection{Multi-Depot Vehicle Routing Problem}
\label{ssec:mdvrp}
%-----------------------------------------------------------------------------%
\textit{Multi-Depot Vehicle Routing Problem (MDVRP)} merupakan salah satu varian dari VRP klasik dimana terdapat lebih dari satu \textit{depot} yang digunakan \cite{montoya-torres_literature_2015}. Gambar \ref{fig:mdvrp-illustration} menunjukkan contoh solusi dari permasalahan VRP yang menggunakan dua depot dan dua buah rute kendaraan yang berasosiasi dengan masing-masing depot. Basically, a solution to this problem is a set of vehicle routes such that: (i) each vehicle route starts and ends at the same depot, (ii) each customer is served exactly once by one vehicle, (iii) the total demand on each route does not exceed vehicle capacity (iv) the maximum route time is satisfied and (v) the total cost is minimized.

\begin{figure}[h]
	\centering
	\includegraphics[width=5cm]{Resources/Images/mdvrp-illustration}
	\caption{Ilustrasi \textit{Multi Depot} VRP}
	\label{fig:mdvrp-illustration}
\end{figure}

The MDVRP can be formalized as follows. Let $G = (V, A)$ be a complete graph, where $V$ is the set of nodes and $A$ is the set of arcs. The nodes are partitioned into two subsets: the customers to be served, $V_C$ = \{1,..., $N$\}, and the multiple depots $V_D$ = \{$N$+1,..., $N+M$\}, with $V_C \cup V_D$ = $V$ and $V_C \cap V_D$ = $\oslash$. There is a non-negative cost $c_ij$ associated with each arc $(i, j) \in A$. The demand of each customer is $d_i$ (there is no demand at the depot nodes). There is also a fleet of $K$ identical vehicles, each with capacity $Q$. The service time at each customer $i$ is $t_i$ while the maximum route duration time is set to $T$. $A$ conversion factor $w_{ij}$ might be needed to transform the cost $c_{ij}$ into time units. In this work, however, the cost is the same as the time and distance units, so $w_{ij}$ = $1$.

In the mathematical formulation that follows, binary variables $x_{ijk}$ are equal to $1$ when vehicle $k$ visits node $j$ immediately after node $i$. Auxiliary variables $y_i$ are also used in the subtour elimination constraints.

\begin{equation}
\label{eq:1}
\begin{aligned}
\sum_{i=1}^{N+M}\sum_{j=1}^{N+M}\sum_{k=1}^{K}c_{ij}x_{ijk};
\end{aligned}
\end{equation}

\begin{equation}
\label{eq:2}
\begin{aligned}
\sum_{i=1}^{N+M}\sum_{k=1}^{K}x_{ijk} = 1  (j=1,..., N);
\end{aligned}
\end{equation}

\begin{equation}
\label{eq:3}
\begin{aligned}
\sum_{j=1}^{N+M}\sum_{k=1}^{K}x_{ijk} = 1  (j=1,..., N);
\end{aligned}
\end{equation}

\begin{equation}
\label{eq:4}
\begin{aligned}
\sum_{i=1}^{N+M}x_{ihk} - \sum_{j=1}^{N+M}x_{hjk} = 0 \\
(k=1,...,K; h=1,...,N+M);
\end{aligned}
\end{equation}

\begin{equation}
\label{eq:5}
\begin{aligned}
\sum_{i=1}^{N+M} \sum_{j=1}^{N+M} d_ix_{ijk} \leq Q (k=1,...,K);
\end{aligned}
\end{equation}

\begin{equation}
\label{eq:6}
\begin{aligned}
\sum_{i=1}^{N+M} \sum_{j=1}^{N+M} (c_{ij}w_{ij} + t_i) x_{ijk} \leq T (k=1,...,K);
\end{aligned}
\end{equation}

\begin{equation}
\label{eq:7}
\begin{aligned}
\sum_{i=N+1}^{N+M} \sum_{j=1}^{N} x_{ijk} \leq 1 (k=1,...,K);
\end{aligned}
\end{equation}

\begin{equation}
\label{eq:8}
\begin{aligned}
\sum_{j=N+1}^{N+M} \sum_{i=1}^{N} x_{ijk} \leq 1 (k=1,...,K);
\end{aligned}
\end{equation}

\begin{equation}
\label{eq:9}
\begin{aligned}
y_i - y_j + (M + N)x_{ijk} \leq N + M - 1; \\
for 1 \leq i \neq j \leq N and 1 \leq k \leq K;
\end{aligned}
\end{equation}

\begin{equation}
\label{eq:10}
\begin{aligned}
x_{ijk} \in \{0, 1\} \forall i,j,k;
\end{aligned}
\end{equation}

\begin{equation}
\label{eq:11}
\begin{aligned}
y_i \in \{0,1\} \forall i;
\end{aligned}
\end{equation}

The objective \ref{eq:1} minimizes the total cost. Constraints \ref{eq:2} and \ref{eq:3} guarantee that each customer is served by exactly one vehicle. Flow conservation is guaranteed through constraint \ref{eq:4}. Vehicle capacity and route duration constraints are found in \ref{eq:5} and \ref{eq:6}, respectively. Constraints \ref{eq:7} and \ref{eq:8} check vehicle availability. Subtour elimination constraints are in \ref{eq:9}. Finally, \ref{eq:10} and \ref{eq:11} define x and y as binary variables.

%-----------------------------------------------------------------------------%
\subsection{Evolution Algorithms}
\label{ssec:evolution-algorithms}
%-----------------------------------------------------------------------------%
\textit{Evolution Algorithms (EAs)} merupakan salah satu \textit{family} dari algoritma optimasi yang terinspirasi dari alam. Pada algoritma evolusi, terdapat proses seleksi dan reproduksi untuk memurnikasi populasi dari kandidat solusi \cite{eugster_many_2003}. Secara umum \textit{lifecycle} dari \textit{Evolution Algorithms} diawali dengan  konfigurasi populasi awal secara random. Kemudian, setiap iterasi dimulai dengan evaluasi atas fungsi obyektif atas individu pada populasi. Berdasarkan hasil evaluasi, kemudian \textit{fitness value} disematkan pada setiap kandidat solusi pada populasi. Kandidat solusi tersebut kemudian mengalami reproduksi dan menghasilkan generasi berikutnya. Proses kemudian berulang dengan menggunakan individu generasi yang baru.

Coevolutionary (CoE) algorithms (CoEA), merupakan pengembangan dari EAs. In standard EAs, evolution is usually viewed as if the population attempts to adapt in a fixed physical environment. In contrast, coevolutionary (CoE) algorithms (CoEA) realize that in natural evolution the physical environment is influenced by other independently-acting biological populations \cite{engelbrecht_coevolution_2007}. CoEA dapat dibagi menjadi dua, tergantung interaksi dari masing-masing spesies: \textit{competitive} dan \textit{cooperative}. Pada \textit{competitive coevolution} setiap individu berkompetisi dengan satu kelompoknya. Sementara pada \textit{cooperative coevolution}, interaksi antar spesies bersifat saling menguntungkan, atau paling tidak, tidak saling membahayakan \cite{engelbrecht_coevolution_2007}.

%-----------------------------------------------------------------------------%
\subsection{Publish/Subscribe Paradigm}
\label{ssec:pub-sub}
%-----------------------------------------------------------------------------%
\textit{Publish/subscribe interaction} merupakan salah satu metode komunikasi antara \textit{client} dan \textit{server}. Paradigma interaksi pada \textit{publish/subscribe} adalah adanya \textit{subscriber} yang memiliki ketertarikan pada suatu \textit{event} atau \textit{pattern of event}, agar dapat dikirimkan notifikasi tentang sebuah \textit{event} oleh \textit{publisher} yang sesuai dengan \textit{interest}-nya \cite{eugster_many_2003}.

Model dasar dari sistem \textit{publish/subscribe}, seperti Gambar \ref{fig:pub-sub-general}, bergantung pada \textit{event notification service} yang menyediakan penyimpanan dan \textit{management of subscription}. \textit{Event service} tersebut berperan sebagai mediator antara \textit{publisher} yang berperan sebagai \textit{producer} dari \textit{event} dan \textit{subscriber} yang berperan sebagai \textit{consumer} dari \textit{event}.

\begin{figure}[h]
	\centering
	\includegraphics[width=8cm]{Resources/Images/pub-sub-general}
	\caption{Arsitektur dasar pada Pub/Sub \cite{eugster_many_2003}}
	\label{fig:pub-sub-general}
\end{figure}

Salah satu keuntungan dari mekanisme publish/subscribe adalah bersifat loose coupling \cite{eugster_many_2003} antara publisher dengan subscriber. Pemisahan informasi (\textit{decoupling}) yang terjadi antara \textit{subscriber} dan \textit{publisher} dapat dipisahkan dalam 3 (tiga) dimensi, yaitu:

\begin{enumerate}
\item \textit{Space decoupling.} \\
Interaksi antara \textit{publisher} dan \textit{subscriber} tidak perlu saling mengetahui satu dengan yang lain. \textit{Publisher} mengirimkan \textit{event}, dan \textit{subscriber} menerima \textit{event} secara tidak langsung melalui \textit{event service}. \textit{Publisher} biasanya tidak memegang referensi terhadap \textit{subscriber}. Begitu juga sebaliknya, \textit{subscriber} biasanya tidak memegang referensi terhadap \textit{publisher}.
\item \textit{Time decoupling.} \\
Pihak-pihak yang berinteraksi, tidak harus berinteraksi pada waktu yang bersamaan. \textit{Publisher} dapat mengirimkan \textit{event} pada saat \textit{subscriber} dalam kondisi terputus. Begitu juga sebaliknya, \textit{subscriber} tetap dapat menerima \textit{event}, meskipun \textit{original publisher} dalam kondisi terputus.
\item \textit{Synchronizing decoupling.} \\
\textit{Publisher} tidak diblok ketika memproduksi \textit{event}, dan juga \textit{subscriber} tetap dapat memperoleh informasi meskipun sedang mengerjakan tugas yang lain. \textit{Publisher} dan \textit{subscriber} tidak berada dalam \textit{main flow}, sehingga tidak berinteraksi secara \textit{synchronous}.
\end{enumerate}

%-----------------------------------------------------------------------------%
\section{Proposed Solution}
\label{sec:proposed-solution}
%-----------------------------------------------------------------------------%
Kelemahan MDVRP ketika digunakan dalam rekomendasi lokasi pencacahan adalah ketiadaan data service time, padahal data itu sangat penting dalam kalkulasi rekomendasi. Untuk itu, MDVRP perlu diintegrasikan dengan mekanisme real-time. Terdapat beberapa mekanisme yang dapat diadopsi pada real-time system, diantaranya: Web service, RPC, message passing, dan publish/subscribe \cite{eugster_many_2003}.

Mekanisme publish/subscribe digunakan dalam penelitian ini karena memiliki sifat \textit{loose coupling}. Selain itu mekanisme publish/subscribe juga sesuai sesuai digunakan pada sistem yang bersifat information driven \cite{muhl_large-scale_2002}. Karena karakteristiknya ini, maka publish/subscribe dapat bekerja secara asynchronous, dimana \textit{request} dan \textit{reply} tidak harus diproses secara berurutan. Pada saat \textit{subscriber} melakukan subscribe, publisher tidak harus dalam kondisi online, begitu juga sebaliknya, pada saat \textit{publisher} melakukan publish, \textit{subscriber} tidak harus dalam kondisi online.

\begin{figure}[h]
	\centering
	\includegraphics[width=8cm]{Resources/Images/system-overview}
	\caption{Garis Besar Sistem}
	\label{fig:system-overview}
\end{figure}

Gambar \ref{fig:system-overview} menunjukkan garis besar sistem, dimana sistem tersusun atas 3 (tiga) komponen utama, yaitu: \textit{publisher}, \textit{subscriber}, dan \textit{message broker}. Komunikasi antara \textit{publisher} dan \textit{subscriber} didasari atas satu kesamaan, yaitu \textbf{event} atau \textbf{topic}. Topik yang digunakan sebagai dasar komunikasi antara publisher dan subscriber dalam penelitian ini adalah \textit{current location} dari masing-masing subscriber.

%-----------------------------------------------------------------------------%
\subsection{Recommendation Publisher}
\label{ssec:recommendation-publisher}
%-----------------------------------------------------------------------------%
VRP Solver diimplementasikan pada sisi publisher, yang akan melakukan pencarian solusi/rute dan kemudian dipublish kepada \textit{subscriber} melalui \textit{message broker}. Sebagaimana sebelumnya disebutkan, bahwasannya \textit{current location} dari masing-masing \textit{subscriber} digunakan sebagai topik. Sebuah thread disiapkan, yang secara periodik akan melakukan pengecekan apakah terdapat topik yang baru pada message broker.

Setiap kali terdapat topik yang baru, sebuah thread akan disiapkan, dan topik tersebut digunakan sebagai ID dari thread yang bersangkutan. Pada thread tersebut terdapat procedure VRP solver yang akan melakukan pencarian solusi/rute. Untuk itu, sebuah threadpool disiapkan untuk menampung thread dari masing-masing topik tersebut secara berurutan berdasarkan kedatangan. Karena topik digunakan sebagai thread ID, maka ukuran dari threadpool akan sama dengan jumlah seluruh lokasi pencacahan.

Setiap thread dalam threadpool kemudian dijalankan secara berurutan, satu session untuk masing-masing thread. Setiap thread yang dijalankan, yang menggunakan topik sebagai ID, akan melakukan pencarian solusi dengan melibatkan seluruh $M$ vehicles dan $(unassigned) N$ locations. Pencarian solusi dilakukan pada procedure VRP solver pada section \ref{ssec:vrp-solver}.

\begin{algorithm}[h]
	\caption{TopicWatcher}
	\begin{algorithmic}[1]
		\renewcommand{\algorithmicrequire}{\textbf{Input:}}
		\renewcommand{\algorithmicensure}{\textbf{Output:}}
		\REQUIRE $C$
		\ENSURE  $T$
		\FOR {$i = 1$ to $len(C)$}
			\FOR {$j = 1$ to $N$}
				\IF {($C_i == E_j$)}
					\STATE $T_j = Thread(C_i, V_1...V_M, (unassigned) E_1...E_N)$
				\ENDIF
			\ENDFOR
		\ENDFOR
		\RETURN $T$
	\end{algorithmic}
\end{algorithm}

Proses dalam procedure VRP solver pada masing-masing thread akan menghasilkan sejumlah minimal $1$ rute, dan maksimal $M$ rute. Setiap rute $R_i$ yang dihasilkan kemudian akan dipublish dengan topik $C_i$. Setelah rute dipublish, \textit{message broker} akan menginformasikan kepada publisher, jumlah \textit{subscriber} yang menerima pesan. Adapun identitas \textit{subscriber} yang menerima pesan tetap tidak diketahui oleh \textit{publisher}.
